Can Auxiliary Indicators Improve COVID-19 Forecasts and Hotspot Prediction?

### 2021-06-13 forecast/paper/paper.tex
Edits through methods D

Define the binary indicator:
$$
Z_{t,\ell} = \indicator{Y^\Delta_{\ell,t} \geq 0.25},
$$
where we use the notation \smash{$Y^\Delta_{\ell,t} = (Y_{\ell,t} - Y_{\ell,
    t-7})/(Y_{\ell,t-7})$}. In other words, $Z_{t,\ell}=1$ if the number of
newly reported cases over the past 7 days has increased by at least 25\%
compared to the preceding week.  When this occurs, we say location $\ell$ is a
\textit{hotspot} at time $t$.

@@

We evaluate forecasts for each forecast date and horizon using the weighted
interval score. Weighted interval score (WIS), a well-known quantile-based
approximation of the commonly-used continuous ranked probability
score~\cite{gneiting-proper-score}. WIS is a proper score, meaning that lower
scores imply better forecasters, and can be thought of as a distributional
generalization of absolute error. The CDC has embraced this metric in the
context of epidemic forecasting and especially for COVID-19~\cite{wis-paper}.
Essentially, pairs of quantile forecasts are treated as a central  interval. For
example, the 0.1 and 0.9 quantiles create an 80\% interval that should, if the
forecaster is well-calibrated, contain the truth about 80\% of the time. A
forecaster whose 80\% interval contains the truth less frequently is overly
confident, while one whose interval contains the truth more frequently (say, by
using an infinitely-wide interval) is failing to take a stand on the future
outcome.

Here, denoting $K$ predicted intervals along with the median  as
$\{\alpha_K/2,\ldots,\alpha_2/2,\  
\alpha_1/2,\ 0.5, 1-\alpha_1/2,\ 1-\alpha_2/2,\ldots, 1-\alpha_K/2\}$, the associated
forecasts as $\{l_K,\ldots, l_2,\ l_1,\ m,\ u_1,\ u_2,\ldots, u_K \}$, and the observed
value (after the forecast was made) as $y$, this metric can be expressed as
\attn{someone please check this} 
\begin{align*}
 \operatorname{WIS}_K &= \frac{1}{K}|m -y | + \frac{2}{K+1}\sum_{k=1}^K 
\frac{1}{\alpha_k}(u_k - l_k)\\ &\quad + \sum_{k=1}^K  
(l_k - y)\indicator{y < l_k} +
(y - u_k)\indicator{y> u_k},
\end{align*}
where $\indicator{a} = 1$ if $a$ is true and $0$ otherwise. Essentially, this
metric decomposes into 4 parts: (1) a penalty for the median, the point forecast
or best guess, being far from the truth; (2) a penalty for the width of each
interval ($u_k - l_k$); (3)  a penalty for overprediction, when the observed
value falls below $l_k$ and hence underneath the interval; and (4) a
corresponding penalty for underprediction. Weighted interval score can also be
written in terms of an average of the quantile or ``pinball'' losses. Setting $\tau_1 =
\alpha_K/2,\ \ldots,\ \tau_{K+1} = 0.5,\ \ldots,\ \tau_{2K+1} = 1-\alpha_K/2$ and
similarly rewriting $q_1 = l_K,\ \ldots,\ q_{K+1}=m,\ \ldots,\ q_{2K+1}=u_K$, then
\begin{align*}
  \operatorname{WIS}_K &= \frac{2}{2K+1} \sum_{k=1}^{2K+1}
                         \left(\indicator{y \leq q_k} -\tau_k\right)
                         (q_k - y).
\end{align*}
While this second expression does not have the convenient decomposition
discussed above, the pinball loss is precisely the objective function that
quantile regression minimizes. So, mathematically, quantile regression should
give parameter estimates that minimize the in-sample weighted interval score.
% DJM: In trying to figure out where 'pinball' comes from, it looks like the
% preprint of the massively cited Koenker and Hallock (2001) JEP article. See
% http://www.econ.uiuc.edu/~roger/research/intro/rq.pdf. They say that tilting
% the absolute value is `pinball logic' which is not a connection I would have
% ever made. The language doesn't seem to have made it to the final version.
% Maybe there's an earlier reference?

When comparing the aggregate performance of the baseline quantile autoregression
without additional signals with 
the indicator-assisted quantile autoregression forecasters, we first
normalize the individual weighted interval scores by the weighted interval
score of a boneheaded model: predicting all future values to be the same as the
most recent observation. The dummy model uses the errors such a forecast has
made in-sample to form the other quantiles. We normalize by 
the dummy model to account for the fact that different locations or
time periods have different magnitudes of COVID-19 incidence (even after
scaling by population). Performing well by this metric means that a particular
forecaster is doing well at all locations and time periods. Without the scaling,
a forecaster that is good at locations and time periods with relatively high
case numbers but poor everywhere else could appear preferable.

For hotspots, we use the area under the curve (AUC) computed by examining the
true-positive and false-negative rates across HRRs at each horizon and forecast
date. That is, for a particular forecast date and a particular horizon, we ask,
how many true hotspots were called hotspots compared to the total number of
hotspots (specificity). The false-negative rate (sensitivity) is the ratio of
correctly classified non-hotspots to the total number of non-hotspots. A perfect
classifier would have both sensitivity and specificity equal to 1. Plotting the
sensitivity on the $y$-axis and $1-$specificity on the $x$-axis where each point
is a forecast date produces an ROC curve for each classifier. We compute the
area under this curve for each forecast horizon to compare the cases-only model
to those adding indicators.

>>>>

Define the binary indicator:
$$
Z_{t,\ell} = \indicator{Y^\Delta_{\ell,t} \geq 0.25},
$$
where we use the notation \smash{$Y^\Delta_{\ell,t} = (Y_{\ell,t} - Y_{\ell,
    t-7})/(Y_{\ell,t-7})$}. In other words, $Z_{t,\ell}=1$ if the number of
newly reported cases over the past 7 days has increased by at least 25\%
compared to the preceding week.  When this occurs, we say location $\ell$ is a
\textit{hotspot} at time $t$.  Empirically, this rule labels about 27\% of
location-time pairs as hotspots, during the prediction period (June 13--December
31, 2020).

@@

For forecasting, we evaluate the probabilistic forecasts produced by the
quantile models \eqref{eq:forecast_ar}, \eqref{eq:forecast_ar_x} using
\textit{weighted interval score} (WIS), which is a well-known quantile-based
scoring rule; see, e.g., \cite{Gneiting:2007}.  WIS is a proper score, which
means that its expectation is minimized by the true quantiles of the target
variable.  The use of WIS in scoring COVID-19 forecasts is discussed by
\cite{Bracher:2021}; and WIS serves as the primary metric for forecast
evaluation in the COVID-19 Forecast Hub.

WIS is typically defined for quantile-based forecasts where the quantile levels
are symmetric around 0.5.  This is the case for our choice in
\eqref{eq:quantile_levels}.  Let $F$ be a forecaster comprised of predicted
quantiles $q_\tau$ parametrized by a quantile level $\tau$.  In the symmetric
case, this is equivalent to a collection of central prediction intervals
$(\ell_\alpha, u_\alpha)$, parametrized by an exclusion probability $\alpha$. 
The WIS of the forecaster $F$, evaluated at the target variable $Y$, is defined
by  
\begin{equation}
\label{eq:wis_intervals}
\WIS(F,Y) = \sum_\alpha \Big\{\alpha(u_\alpha - \ell_\alpha) + 2 \cdot
\dist(Y, [\ell_\alpha, u_\alpha])\Big\},  
\end{equation}
where $\dist(a, S)$ is the distance between a point $a$ and set $S$ (the
smallest distance between $a$ and an element of $S$).  Note that, corresponding 
to \eqref{eq:quantile_levels}, the exclusion probabilities are $\alpha \in
\{0.05, 0.2, 0.5, 1\}$, resulting in 4 terms in the above sum.  By 
straightforward algebra, it is not hard to see WIS has an alternative 
representation in terms of the predicted quantiles themselves,
\begin{equation}
\label{eq:wis_quantiles}
\WIS(F,Y) = 2 \sum_\tau \phi_\tau(Y - q_\tau), 
\end{equation}
where $\phi_\tau(x) = \tau |x|$ for $x \geq 0$ and $\phi_\tau(x) = (1-\tau)
|x|$ for $x<0$, which is often called the ``tilted absolute'' loss.  While
\eqref{eq:wis_quantiles} is more general (it can accomodate asymmetric quantile
levels), the first definition \eqref{eq:wis_intervals} is usually preferred in
presentation, because the score nicely decouples into a ``sharpness'' component
(first term in each summand) and ``under/overprediction'' component (second term
in each summand).  The second form given in \eqref{eq:wis_quantiles} is
especially noteworthy in our current study as it reveals WIS is the same as the
quantile regression loss that we use to train our forecast models (in other
words, our models are fit to optimize WIS average over the training set).

For hotspot prediction, we evaluate the probabilistic classifiers produced by
the logistic models \eqref{eq:hotspot_ar}, \eqref{eq:hotspot_ar_x} using the
area under the curve (AUC) of their true positive versus false positive rate
curve (which is traced out by varying the discrimination threshold).
